---
title: "Assignment1_607"
author: "Vishal Arora"
date: "Feburary 1, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#DATA 607 Week 1 Assignment - Loading Data into a Data Frame

# Introduction 

In week 1 assignment for DATA607. we will load data into dataframe that is provided by UCI Mushroom dataset located : https://archive.ics.uci.edu/ml/datasets/Mushroom

# About The data 

This data set includes descriptions of hypothetical samples corresponding to 23 species of gilled mushrooms in the Agaricus and Lepiota Family .  Each species is identified as definitely edible, definitely poisonous, or of unknown edibility and not recommended.

#Data Dictionary :-
 
Attribute Information:

0. Type     : Edible e , Poisonous = p

1. cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s 

2. cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s 

3. cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y 

4. bruises: bruises=t,no=f 

5. odor: almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s 

6. gill-attachment: attached=a,descending=d,free=f,notched=n 

7. gill-spacing: close=c,crowded=w,distant=d 

8. gill-size: broad=b,narrow=n 

9. gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e, white=w,yellow=y 

10. stalk-shape: enlarging=e,tapering=t 

11. stalk-root: bulbous=b,club=c,cup=u,equal=e, rhizomorphs=z,rooted=r,missing=? 

12. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s 

13. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s 

14. stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y 

15. stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y 

16. veil-type: partial=p,universal=u 

17. veil-color: brown=n,orange=o,white=w,yellow=y 

18. ring-number: none=n,one=o,two=t 

19. ring-type: cobwebby=c,evanescent=e,flaring=f,large=l, none=n,pendant=p,sheathing=s,zone=z 

20. spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r, orange=o,purple=u,white=w,yellow=y 

21. population: abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y 

22. habitat: grasses=g,leaves=l,meadows=m,paths=p, urban=u,waste=w,woods=d


#Problem Statement:-
Load data from given URL(https://archive.ics.uci.edu/ml/datasets/Mushroom) into R , subset and create a new Data frame selecting few columns from original including 1st column. Provide meangingfull name to columns Headers, and also update the values of each column based on the data dictionary values.

##First step to load the necessary libraries required for this assignment.
```{r}
library(stringr)
library(XML)
library(maps)
library(httr)
```


##R-Code starting :-
```{r}

mush_table <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data",     header=FALSE , sep="," )

mushrooms <- as.data.frame(mush_table)


# subsetting the data to create a a new data frame with 5 columns
mush_subset <- subset(mushrooms, select=c(1,2,3,4,6,22,23))

#Giving columns meaningfull names
colnames(mush_subset) <- c("Type","Shape","Surface","Color","Odor","Population_Type","Habitat")

#Changing the column values with meaningfull values based on conditional statements
mush_subset$Type <- ifelse(str_detect(mush_subset$Type, "e") == TRUE, "Edible", "Poisonous")

#bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s 
mush_subset$Shape <- ifelse(str_detect(mush_subset$Shape, "x") == TRUE, "convex", ifelse(str_detect(mush_subset$Shape, "c") == TRUE, "conical",ifelse(str_detect(mush_subset$Shape, "b") == TRUE, "bell" , ifelse(str_detect(mush_subset$Shape, "k") == TRUE, "knobbed",ifelse(str_detect(mush_subset$Shape, "f") == TRUE, "flat","sunken")))))

#fibrous=f,grooves=g,scaly=y,smooth=s
mush_subset$Surface <- ifelse(str_detect(mush_subset$Surface, "f") == TRUE, "fibrous", ifelse(str_detect(mush_subset$Surface, "g") == TRUE, "grooves",ifelse(str_detect(mush_subset$Surface, "y") == TRUE, "scaly" , "smooth")))

#brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y 
mush_subset$Color <- ifelse(str_detect(mush_subset$Color, "n") == TRUE, "brown", ifelse(str_detect(mush_subset$Color, "b") == TRUE, "buff",ifelse(str_detect(mush_subset$Color, "c") == TRUE, "cinnamon" , ifelse(str_detect(mush_subset$Color, "g") == TRUE, "gray",ifelse(str_detect(mush_subset$Color, "p") == TRUE, "pink",ifelse(str_detect(mush_subset$Color, "u") == TRUE, "purple",ifelse(str_detect(mush_subset$Color, "r") == TRUE, "red",ifelse(str_detect(mush_subset$Color, "w") == TRUE, "white","yellow"))))))))

#almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s
mush_subset$Odor <- ifelse(str_detect(mush_subset$Odor, "a") == TRUE, "almond", ifelse(str_detect(mush_subset$Odor, "l") == TRUE, "anise",ifelse(str_detect(mush_subset$Odor, "c") == TRUE, "creosote",ifelse(str_detect(mush_subset$Odor, "y") == TRUE, "fishy" , ifelse(str_detect(mush_subset$Odor, "f") == TRUE, "foul",ifelse(str_detect(mush_subset$Odor, "m") == TRUE, "musty",ifelse(str_detect(mush_subset$Odor, "p") == TRUE, "pungent",ifelse(str_detect(mush_subset$Odor, "s") == TRUE, "spicy","none"))))))))

#population: abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y 
mush_subset$Population_Type <- ifelse(str_detect(mush_subset$Population_Type, "a") == TRUE, "abundant", ifelse(str_detect(mush_subset$Population_Type, "c") == TRUE, "clustered",ifelse(str_detect(mush_subset$Population_Type, "n") == TRUE, "numerous" , ifelse(str_detect(mush_subset$Population_Type, "s") == TRUE, "scattered",ifelse(str_detect(mush_subset$Population_Type, "v") == TRUE, "several","solitary")))))



#habitat: grasses=g,leaves=l,meadows=m,paths=p, urban=u,waste=w,woods=d
mush_subset$Habitat <- ifelse(str_detect(mush_subset$Habitat, "g") == TRUE, "grasses", ifelse(str_detect(mush_subset$Habitat, "l") == TRUE, "leaves",ifelse(str_detect(mush_subset$Habitat, "m") == TRUE, "meadows" , ifelse(str_detect(mush_subset$Odor, "p") == TRUE, "paths",ifelse(str_detect(mush_subset$Habitat, "u") == TRUE, "urban",ifelse(str_detect(mush_subset$w, "p") == TRUE, "waste","woods"))))))



```


## Summary:-
The data frame after subsetting, has been given meaningfull name and updating column values is as below.
```{r}
head(mush_subset,n=20)
```





